Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning

### 2019-11-16 xrb2/paper.tex
paper II stub

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Dynamics of Laterally Propagating Flames in X-ray Bursts}

\shorttitle{Lateral Flame Dynamics}
\shortauthors{Eiden et al.}

@@

\correspondingauthor{Michael Zingale}
\email{Michael.Zingale@stonybrook.edu}

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
We investigate the structure of laterally-propagating flames through
the highly-stratified burning layer in an X-ray burst.  Simple \MarginPar{is simple the right word here?} \MarginPar{It might be better to say simulations are performed through a simplified atmosphere?}
two-dimensional simulations of flame propagation are performed through
a rotating plane-parallel atmosphere, exploring the structure of the
flame.  We discuss the approximations needed to capture the length and
time scales at play in an X-ray burst.  Our studies complement 
other multidimensional studies of burning in X-ray bursts.
\end{abstract}

@@

X-ray bursts (XRBs) are thermonuclear explosions in an accreted H or
He layer on the surface of a neutron star (see~\citealt{galloway:2017}
for a review).  Observations of bursts can assist in constraining the
properties of the underlying neutron star, helping to illuminate the
nuclear equation of state~\citep{steiner:2010,ozel2016masses}.  Extensive
observations of brightness oscillations during the rise (the initial phase of the burst where the observed flux rapidly increases) have provided
evidence that the burning begins in a localized region and spreads
over the surface of the neutron
star~\citep{bhattacharyya:2006,bhattacharyya:2007,chakraborty:2014}.

One dimensional studies of XRBs have been very successful in predicting the
lightcurves and recurrence times (see, e.g., \citealt{woosley-xrb}).
These assume spherical symmetry and thus cannot capture the effects of
localized burning spreading across the neutron star.  These studies
have also been used to explore the sensitivity of the burst
observables to accretion and reaction
rates~\citep{cyburt:2010,Jose2010a,Lampe2016}, and to model individual
bursts~\citep{johnston:2019}.

Multidimensional simulations of burning on a neutron star are more
difficult, with both the temporal and spatial scales presenting
challenges (see \citealt{astronum_2018} for an overview).  For the spatial scales, we need to
resolve the reaction zone, $\mathcal{O}(10~\mbox{cm})$ or smaller, the
scale height of the atmosphere, $\mathcal{O}(500~\mbox{cm})$, and the
Rossby scale where the Coriolis force balances the lateral pressure
gradient, $\mathcal{O}(10^5~\mbox{cm})$~\citep{spitkovsky2002}.  For
the temporal scales, capturing the rise, $\mathcal{O}(1~\mbox{s})$,
and the decay of the burst, $\mathcal{O}(10~\mbox{s})$, as well as the
accretion period between bursts, $\mathcal{O}(10^4~\mbox{s})$, is
currently beyond the ability of multidimensional hydro codes.
Nevertheless, significant progress has been made in understanding the
multidimensional nature of XRBs, through various approximations.

Laterally propagating detonations were modeled by
\citet{fryxellwoosley82,hedet}. However, since it is difficult to detonate helium at the
densities found in normal XRBs, and even harder to detonate hydrogen
because of the waiting times for weak reactions, these would only occur at
very high densities.  This means that
detonations may only really be applicable to superbursts (where carbon
is the reactant)~\citep{Weinberg2006b,Weinberg2007}.

Global multidimensional studies were performed by
\citet{spitkovsky2002}, where it was demonstrated that the Coriolis
force plays an important role in confining the burning as it spreads
across the neutron star surface.  These calculations used the
shallow-water approximation, so the vertical details of the
atmosphere's structure we not captured.  Their model showed that the
horizontal pressure gradient between the ash and fuel can be important
in accelerating the burning front.

Small-domain studies of convective burning in XRBs preceding flame
development have been done in two-dimensions \citep{lin:2006,xrb,xrb2}
and three-dimensions \citep{xrb3d}.  These calculations used low Mach
number methods, which approximate the hydrodynamics equations to filter
soundwaves, enabling large timesteps and efficient modeling of
subsonic convection.  While these calculations could not support
the lateral differences needed for flame spreading, they can help
understand the role that convection plays in distributing the initial
burning products vertically throughout the neutron star atmosphere as well
as the nature of any turbulence the burning front might encounter as it propagates through the atmosphere.

\cite{cavecchi:2013,art-2015-cavecchi-etal,art-2016-cavecchi-etal,Cavecchi2017,Cavecchi2019} \MarginPar{need a summary}

For deflagrations, we either need to resolve the structure of the
reaction zone or use a flame model.  Flame models usually assume that
the flame structure is thin compared to the size of the system (see,
e.g., \citet{Ropke2007} for applications to Type Ia supernovae).  For
XRBs however, the flame thickness is comparable to the scale height of
the atmosphere, so we cannot use these approximations.  Accurate
models of flames in XRBs therefore require that we resolve the thermal width,
which is $\mathcal{O}(10~\mbox{cm})$ for helium flames~\citep{Timmes00}.

The goal of this study is to understand what numerical and physical
approximations are required to perform a full hydrodynamical,
multidimensional simulation of flame propagation through the
atmosphere of a neutron star.  For simplicity in this first set of
calculations, we will use a pure helium composition.  These studies
complement the prior multidimensional studies described above in helping
us to build a picture of the dynamics of X-ray bursts.

@@

All simulations are performed with the \castro\ hydrodynamics
code~(\citealt{castro}; see also \citealt{astronum:2017} for a recent
description).  We evolve the system of fully compressible Euler
equations for reacting flow:
\begin{eqnarray}
\frac{\partial( \rho X_k)}{\partial t} &=& - \nabla\cdot (\rho \Ub X_k) + \rho \omegadot_k \\
%
\frac{\partial (\rho \Ub)}{\partial t} &=& - \nabla\cdot (\rho \Ub \Ub) - \nabla p +
    \rho \gb \nonumber \\
  &&-2 \rho \Omegab\times \Ub - \Omegab \times (\Omegab \times \rb) \\
%
\frac{\partial (\rho E)}{\partial t} &=& - \nabla \cdot (\rho \Ub E + p\Ub) +
    \nabla \cdot \kth \nabla T + \rho \epsdot +\nonumber\\
  && \rho \Ub \cdot \gb - \rho (\Omegab \cdot \rb)(\Omegab \cdot \Ub) + \rho |\Omegab|^2 (\Ub \cdot \rb)
\end{eqnarray}
Here, $\rho$ is the mass density, $\Ub$ is the velocity, $p$ is the pressure and $E$ is
the specific total energy, which is related to the specific internal
energy as $e = E - |\Ub|^2/2$.  Species are described by mass
fractions, $X_k$ (such that $\sum_k X_k = 1$), and creation rates,
$\omegadot$, and are related to the total specific energy generation
rate, $\epsdot$.  The total mass conservation,
\begin{equation}
\frac{\partial \rho}{\partial t} = - \nabla\cdot (\rho \Ub)
\end{equation}
implies $\sum_k \omegadot_k = 0$.  \castro\ uses an unsplit piecewise
parabolic method (PPM) with characteristic tracing for solving the
hydrodynamics~\citep{ppm,millercolella:2002}, generalized to an
arbitrary equation of state~\citep{zingalekatz}.  Reactions are
incorporated via Strang splitting~\citep{strang:1968}, giving a method
that is overall second-order accurate in space and time.
\castro\ uses the \amrex\ adaptive mesh refinement
library~\citep{amrex_joss} to manage a hierarchy of grids at different
resolutions.

An equation of state of the form $p = p(\rho, e, X_k)$ completes the
thermodynamic description of the system.  Finally, $\kth$ is the
thermal conductivity and $T$ is the temperature.  Diffusion is treated
explicitly in time. \MarginPar{Have not defined what r is in eqs 1-3, and Omega and g are only defined with respect to eq. 5}

Since the neutron star rotates, we work in a corotating frame, with
angular velocity $\Omegab$.  Further, for the two-dimensional
simulations presented here, we work in axisymmetric coordinates,
but we advect a third component of velocity, coming out of the
simulation plane, that participates in the Coriolis force (sometimes
described as a 2.5D simulation).  We will take the
\castro\ $x$-coordinate to be the cylindrical radial coordinate with
corresponding velocity $u$, the \castro\ $y$-coordinate to be the
cylindrical vertical coordinate with corresponding velocity $v$, and the
\castro\ $z$-coordinate to be the cylindrical azimuthal coordinate,
with corresponding velocity $w$.  A righthanded coordinate system has
positive $w$ pointing out of the page.  We will take $\Omegab =
\Omega_0 \hat{\bf y}$ for the angular rotation rate, and $\gb = -g
\hat{\bf y}$ for the gravitational acceleration, with $g$ constant.
With these choices, the Coriolis force is:
\begin{equation}
-2\rho \Omegab \times \Ub =
   -2\rho \left ( \Omega_0 w \hat{x} - \Omega_0 u \hat{z} \right )
\end{equation}
We will neglect the centrifugal force---with our plane-parallel
geometry, this will act only in the lateral direction, and is not
expected to greatly affect the dynamics.  Carrying the Coriolis force
allows us to capture the geostrophic balance\MarginPar{ref, Spitkovsky 2002?} that sets
up via lateral hydrostatic equilibrium.  In the discussions below, we will use the Castro coordinate
names, $(x, y)$ in our notation.

Writing the momentum equation in terms of the $u$, $v$, and $w$
components, and neglecting the centrifugal force, we have:
\begin{align}
\frac{\partial (\rho u)}{\partial t} + \nabla \cdot (\rho u \Ub) +
     \frac{\partial p}{\partial x} &= -2\rho \Omega_0 w  \\
%
\frac{\partial (\rho v)}{\partial t} + \nabla \cdot (\rho v \Ub) +
     \frac{\partial p}{\partial y} &= -\rho g \\
%
\frac{\partial (\rho w)}{\partial t} + \nabla \cdot (\rho w \Ub) +
     \cancelto{0}{\frac{\partial p}{\partial z}} &=
    2\rho \Omega_0 u
\end{align}
where we cancel $\partial p/\partial z$ because there are no
variations in the azimuthal direction.  This allows us to recast the
$w$-velocity equation as a simple advection equation:
\begin{equation}
\frac{\partial w}{\partial t} + \Ub \cdot \nabla w = 2\Omega_0 u
\end{equation}
In our geometry, the flame will propagate from left to right, so $u$
will be positive and the Coriolis force results in $w > 0$ (out of the
simulation plane).  The algorithmic implementation of rotation in
\castro\ is described in \cite{wdmergerI}.

We use a general stellar equation of state with nuclei (treated as an
ideal gas), photons, and degenerate/relativistic electrons, as
described in \cite{timmes_swesty:2000}.  To model reactions, we use a 
13-isotope alpha chain, derived from the {\tt aprox13}
network~\citep{timmes_aprox13}.  For one of our runs, we use the
smaller 7-isotope network described in \citet{iso7}.  We integrate the
network using the VODE integration package~\citep{vode}, and our
implementation is provided in the StarKiller Microphysics
source~\citep{starkiller}.
We note that we do not explicitly model viscosity.  The reactions will
provide the smallscale cutoff to the instabilities and turbulence at
the flame front.  We also do not include species
diffusion---astrophysical flames tend to have large Lewis numbers, so
this is not expected to be important~\citep{timmeswoosley:1992}.
Finally, we use the thermal conductivities described in
\citet{Timmes00}.

All simulations use adaptive mesh refinement to refine on the
atmosphere (leaving the space between the top of the atmosphere and
upper boundary at low resolution). As seen in Figure \ref{fig:grids}, 
we use up to 3 refinement levels in addition to the base grid, the first
one a factor of 4 finer than the previous and the remaining each a
factor of 2 finer than the previous. \castro\ subcycles in time, so the 
finer grids are evolved with a finer timestep than the coarse grids. 
Occasionally, the timestep chosen at the start of a cycle will violate
the CFL condition during the advancement of the finer grids. In this
case, we restart the finer grid evolution with a smaller timestep,
subcycling within the larger timestep hierarchy.  We use a CFL number
of 0.8 for our simulations.  
% \MarginPar{true for all?} 
The base grid for our standard
simulations is $768\times 192$ zones, and the equivalent finest grid
when 3 refinement levels are added would be $12288\times 3072$ zones.
Our standard domain is $1.2288\times 10^5~\mathrm{cm} \times
3.072\times 10^4~\mathrm{cm}$, corresponding to $10~\mathrm{cm}$ resolution
on the finest grid.  We only refine the fuel layer in the left half of
the domain at the highest resolution (and only down to densities of
$2.5\times 10^4~\gcc$), since this is where we expect the flame to
propagate.  At the start of the simulation, $3.4\%$ of the domain is
at the finest resolution.  This increases to $7.4\%$ by the end of the
simulation, because of the increase in the scale height of the atmosphere
behind the flame.

\begin{figure}[t]
	\plotone{grids.pdf}
	\caption{\label{fig:grids}A slice through part of a 2D simulation showing the four-level
		grid structure. Note that the boxes shown are not the simulation zones, which span
		$10~\mathrm{cm}$ at the finest level, but subdomains containing approximately equal
		numbers of zones that are each managed by a separate MPI process. A coarse base grid
		extending to the upper boundary is drawn in white, with the green, blue, and black
		grids showing the jumps in refinement needed to fully resolve the fuel layer and
		underlying neutron star.}
\end{figure}

Thermal diffusion is modeled explicitly, using a predictor-corrector
scheme to achieve second-order accuracy.  A verification test of the
diffusion scheme is shown in Appendix~\ref{app:diffusion}.  The
explicit thermal diffusion requires a timestep limiter of the form:
\begin{equation}
\Delta t_\mathrm{diff} \le \frac{1}{2} \frac{\Delta x^2}{\mathcal{D}}
\end{equation}
where $\mathcal{D} = \kth/(\rho c_v)$ is the thermal diffusivity.  The
diffusivity increases rapidly at the top of the atmosphere, causing
these low density regions to determine the overall timestep for the
simulations.  Therefore, we disable thermal conduction at low
densities where it is not expected to be important.

We use hydrostatic boundary conditions on the lower boundary, using
a discretized hydrostatic equilibrium equation of the form:
\begin{equation}
\label{eq:hse}
p_i = p_{i-1} + \frac{1}{2} \Delta y (\rho_i + \rho_{i-1}) g
\end{equation}
and holding the temperature constant in the ghost cells.  This is solved
together with the equation of state.  The velocity is reflected at this boundary.
This procedure follows the form described in \citet{ppm-hse}.  The
left boundary is reflecting and the right boundary is a zero-gradient
outflow.  The top boundary sets the state to simply the conditions
in our outer buffer region of the initial model (see below), with the
normal velocity set to the larger of zero or the velocity at the top
of the domain (this prevents incoming velocities at the top) and the
transverse velocities set to zero.

When we begin the simulation, there is a transient phase as the flame
gets established.  Material that is forced upward will encounter the
steep density gradient at the top of the atmosphere and accelerate as
it is blown out of the atmosphere.  Eventually this material will fall
back to the top of the atmosphere.  In this paper, we are mostly
concerned with the behavior of the flame and not any material that is
violently blown out of the top of the atmosphere, so we apply a sponge
to this region.  This is similar to the method we previously used in \cite{xrb3d}.
The sponge drives the velocity of the material in the low
density \MarginPar{add mathematical form of sponge} regions at the top
of the atmosphere to zero.  This sponging helps increase our timestep
as well.

%======================================================================
% Flame
%======================================================================
\section{Flame Properties}\label{Sec:Flame}

The speed and thickness of a laminar helium flame are determined by the
energy generation rate and conductivity, and scale roughly as 
\begin{equation}
v_f \approx \sqrt{k_\mathrm{th} \dot{\epsilon}} \qquad
\lambda_f \approx \sqrt{\frac{k_\mathrm{th}}{\dot{\epsilon}}}
\end{equation}
\citep{orourke:1979,khokhlov:1993}.
At the densities we consider in this simulation, the pure He flame
speed is quite slow and would require a prohibitively long integration
time to see significant evolution of the burning.  We therefore
consider boosted flames in this first paper, to accelerate the burning
and allow us to understand the qualitative effects of laterally
propagating flames.  To boost the flame while keeping the thickness
the same, we can multiply both the burning rate and the conductivity
by the same factor.  For our standard calculations, we choose 10 for
each, to give a 10$\times$ faster flame speed.  We call this the ``10/10''
flame.  We will also do a simulation with the reactions and
conductivity both boosted by 5, the ``5/5'' flame.

To understand the time and length scales involved in flame
propagation, we do a 1D simulation of a laminar flame using our
microphysics.  Figure~\ref{fig:flame} shows the flame properties for a 
% \AssignTo{Kiran and Mike will remake with the 10/10 boosted flame}
He flame using the standard conductivities and our simple reaction \MarginPar{standard?  this is a boosted flame, right?}
network.  This flame had a density of $2\times 10^6~\gcc$ and 
% \MarginPar{double check}
temperature of $5\times 10^7$~K.  We observe that this flame speed is
about $10^5~\cms$, the flame width is about 40~cm, and it
takes about 3~ms to settle into a sustained flame.  Note that this speed
is quite small compared to the speed estimated in
\citet{spitkovsky2002}.  We expect a multidimensional flame to
accelerate due to hydrodynamics interactions (wrinkling, turbulence
interactions, directed flows feeding fuel into the flame, ...).

\begin{figure*}[t]
\plottwo{flame.pdf}{speed.pdf}
\caption{\label{fig:flame} Time-evolution of the 10$\times$ boosted 1D
	laminar flame. The left plot shows temperature and nuclear energy
	generation profiles at 11 different time points, while the right plot
	shows flame propagation speed and flame thickness as functions of time.}
\end{figure*}

We measure the flame width as:
\begin{equation}
\lambda_f \equiv \frac{\Delta T}{\max\{|\nabla T|\}}
\end{equation}
Experience with modeling resolved flames suggests that we need a
spatial resolution, $\Delta x$ of $\lambda_f/\Delta x \sim 5$~\citep{SNld}.  These
conditions represent the bottom of the He layer.  As the density
decreases with altitude, the flame thickness increases and the flame
speed decreases, so we will easily resolve the flame structure throughout
the rest of the atmosphere.

Flame propagation rates from both the 1D laminar flames and full 2D simulations are given in Table \ref{table:flame_speeds}.  \MarginPar{I think we should only report the laminar speeds here and defer the turbulent speeds until the section where we talk about how they are computed}

\begin{deluxetable}{lcc}
	\tablecaption{\label{table:flame_speeds}Turbulent and laminar flame speeds.}
	\tablehead{\colhead{run} & \colhead{$s_T$ (km s$^{-1}$)} & \colhead{$s_L$ (km s$^{-1}$)}}
	\startdata
	10/10 boost & $9.18 \pm 0.03$  & $1.06 \pm 0.01$ \\
	5/5 boost & $4.00 \pm 0.01$ & $0.56 \pm 0.01$ \\
	iso7 network & $7.56 \pm 0.02$ & $-$ \\
	lowres & $9.33 \pm 0.04$ & $-$ \\
	1000 Hz & $18.6 \pm 0.16$ & $-$ \\
	\enddata
	%
\end{deluxetable}

%======================================================================
% Initial Model
%======================================================================
\section{Initial Model}\label{Sec:inital_model}

% \begin{figure}[t]
% \centering
% \plotone{model}
% \caption{\label{fig:sketch} A schematic of our model setup.}
% \end{figure}

We wish to create an initial atmosphere consisting of a hot
``post-flame'' region and a cooler atmosphere that the flame will
laterally propagate into.  We put the hot region at the very left of
the domain (the origin of the axisymmetric coordinates).  To create
these initial conditions, we produce two different hydrostatic models,
a ``hot'' model that will represent the perturbation that drives the
flame and a ``cool'' model that will represent the state ahead of the
flame.  These will have different scale heights.  To create these
models, we break the vertical structure of the atmosphere into 4
layers: the underlying neutron star, a ramp-up to the base of the
accreted atmosphere, a fuel layer representing the bulk of the
atmosphere where the flame will propagate, and an outer, low density,
isothermal buffer above the atmosphere that allows us
to capture expansion and explosive dynamics.  
% This is sketched in
% Figure~\ref{fig:sketch}.

The temperature profile in the star and ramp region is given as:
\begin{equation}
T = T_\star + \frac{1}{2} (T_\mathrm{hi} - T_\star) \left [ 1 + \tanh\left( \frac{\tilde{y}}{2 \delta_\mathrm{atm}} \right ) \right ]
\end{equation}
with
\begin{equation}
\tilde{y} = y - H_\star - \frac{3}{2} \delta_\mathrm{atm}
\end{equation}
Here, $\delta_\mathrm{atm}$ is a characteristic width of the
transition ramp, $T_\star$ is the temperature of the neutron star, and
$T_\mathrm{hi}$ is the highest temperature in the HSE model---it will
represent the base of the fuel layer.

The species mass fractions use this same profile, switching from a set
describing the underlying star, ${X_k}_\star$, and the set for the
accreted material, ${X_k}_\mathrm{atm}$, which is used in the
isentropic and outer regions.  Note, since the profile above is
linear in $X$, if the initial mass fractions sum to one, then the blended
mass fractions in the ramp region also sum to one.

We specify the density, $\rho_\mathrm{int} = \rho(y = H_\star)$ as the
starting point for the integration of hydrostatic equilibrium.  This
is just below the ramp-up region---this ensures that regardless of
what the peak temperature ($T_\mathrm{hi}$) is, the state beneath the
ramp-up region remains unchanged.  Therefore, we will still be in
lateral equilibrium in the star region.  We will denote the density
where $T = T_\mathrm{hi}$ as $\rho_\mathrm{fuel}$.

Creating the model involves specifying $T_\star$, $T_\mathrm{hi}$,
$T_\mathrm{lo}$, $\rho_\mathrm{int}$, $H_\star$,
$\delta_\mathrm{atm}$, ${X_k}_\star$, and ${X_k}_\mathrm{atm}$.  We
then integrate outwards from the base of the ramp region ($y =
H_\star$), enforcing the discrete form of hydrostatic equilibrium,
Eq.~\ref{eq:hse}.  Integrating upwards, we find $p_i$ and
$\rho_i$ using a Newton solve together with the equation of state, with either the temperature
specified, $T_i = T(p_i, \rho_i, \{X_k\}_i)$ (in the isothermal, ramp,
and buffer layers) or constant entropy, $s_i = s(p_i, \rho_i,
\{X_k\}_i)$, in the fuel layer.  This follows the procedures described in \citet{ppm-hse}.  We
use a constant temperature for all $y < H_\star +
3\delta_\mathrm{atm}$.  Above this, we switch to an isentropic atmosphere until the
temperature drops to a floor value, $T_\mathrm{lo}$, at which point we
again keep the temperature constant.  The integration of the
atmosphere continues until the density falls to a low density cutoff,
$\rho_\mathrm{cutoff}$.  The material above this height is taken to
have constant density and temperature.

\begin{deluxetable}{lcc}
\tablecaption{\label{table:params} Initial model parameters.}
\tablehead{\colhead{parameter} & \colhead{cool} & \colhead{hot}}
\startdata
$T_\star$       & \multicolumn{2}{c}{$10^8$~K} \\
$T_\mathrm{hi}$ & $2\times 10^8$~K & $1.4\times 10^9$~K \\
$T_\mathrm{lo}$ & \multicolumn{2}{c}{$8\times 10^6$~K} \\
$\rho_\mathrm{int}$ & \multicolumn{2}{c}{$3.43\times 10^6~\gcc$} \\
$\rho_\mathrm{fuel}$\tablenotemark{a} & $2.36\times 10^6~\gcc$ & $1.20\times 10^6~\gcc$ \\
$\rho_\mathrm{cutoff}$ & \multicolumn{2}{c}{$10^{-4}~\gcc$} \\
$g$             & \multicolumn{2}{c}{$-1.5\times 10^{14}~\mathrm{cm~s^{-2}}$} \\
$H_\star$       & \multicolumn{2}{c}{2000~cm} \\
$\delta_\mathrm{atm}$ & \multicolumn{2}{c}{50~cm} \\
$X_\star(\isotm{Ni}{56})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
$X_\mathrm{atm}(\isotm{He}{4})\tablenotemark{b}$ & \multicolumn{2}{c}{1.0} \\
\enddata
%
\tablenotetext{a}{This is not an input parameter, but instead is
  computed during integration.  We list it here for reference.}
%
\tablenotetext{b}{All other species are taken as 0.}
\end{deluxetable}

The choice of factors in front of $\delta_\mathrm{atm}$ were designed
to make sure the peak $T$ is attained at the desired density
of the burning layer.

We create two models, a ``cool'' model representing the atmosphere
ahead of the flame and a ``hot'' (or perturbed) model that will
initiate the flame by driving the burning.  \MarginPar{This has already been said at the start of this section} The parameters we use for
the model generation are listed in Table~\ref{table:params}. \MarginPar{reference and discuss the figure of initial model here}

We blend the hot and cold models laterally to produce the perturbation
needed to initiate a localized flame, with the hot model at the
origin of the axisymmetric geometry.  The blending is done as:
\begin{align}
p(x,y) &= f(x) p_\mathrm{hot}(y) + [1-f(x)] p_\mathrm{cool}(y) \\
\rho(x,y) &= f(x) \rho_\mathrm{hot}(y) + [1-f(x)] \rho_\mathrm{cool}(y) \\
X_k(x,y) &= f(x) {X_k}_\mathrm{hot}(y) + [1-f(x)] {X_k}_\mathrm{cool}(y)
\end{align}
with
\begin{equation}
f(x) = \begin{cases}
     1 & x < x_\mathrm{pert} \\
   1 - \frac{x - x_\mathrm{pert}}{\delta_\mathrm{blend}} & x_\mathrm{pert} \le x \le x_\mathrm{pert} + \delta_\mathrm{blend} \\
     0 & x > x_\mathrm{pert} + \delta_\mathrm{blend}
\end{cases}
\end{equation}
Since the equation of hydrostatic equilibrium is linear and our
blending is a linear combination of two models in hydrostatic
equilibrum, the blended model is also in vertical equilibrium initially.  We choose
$x_\mathrm{pert} = 1.024\times 10^4$~cm and $\delta_\mathrm{blend} = 2048$~cm.  Once \MarginPar{double check these numbers}
the blended model is constructed, we compute $T(x,y)$ and $(\rho e)(x,y)$
from the equation of state,
\begin{align}
  T(x,y) &= T(\rho(x,y), p(x,y), X_k(x,y)) \\
  (\rho e)(x,y) &= \rho(x, y) \cdot e(\rho(x,y), p(x,y), X_k(x,y)) 
\end{align}

\begin{figure}[t]
\centering
\epsscale{0.75}
\plotone{initial_model_paper}
\caption{\label{fig:initial_models} Our ``cool'' (solid) and ``hot''
  (dashed) initial models, showing both the density and temperature.}
\epsscale{1.0}
\end{figure}

The initial model is created on a uniform grid at the resolution
corresponding to the finest level of refinement.  At regions of the
atmosphere that are not refined to the finest level, we interpolate
density, pressure, and composition on the grid and then obtain the
temperature from the EOS.

@@

Table~\ref{table:sim_names} summarizes the simulations that we ran. \MarginPar{This table has different model names to the previous table}
The majority of them used a reaction rate boosting of $10$ and a
conductivity boosting of $10$, which should increase the flame speed
by a factor of $10$.  Most simulations used a resolution of
$10~\mbox{cm}$ and a domain width of a little more than one kilometer.  
We use an artifically high rotation rate of
$2000~\mbox{Hz}$, which gives a Rossby length of
\begin{equation}
L_R = \frac{\sqrt{g H_0}}{\Omega} \sim 3\times 10^4~\mathrm{cm}
\end{equation}
using a scale height $H_0 = 10^3~\mathrm{cm}$.  This is about one
quarter of the domain width.  The run at $20~\mbox{cm}$ resolution used
one fewer level of refinement.  The slower rotating case using a
slightly wider domain to accommodate the expected larger Rossby
length.  We note that the entire simulation framework for these
calculations is freely available in the \castro\ github
repository\footnote{\url{https://github.com/AMReX-Astro/Castro}, using
  the {\tt flame\_wave} setup.}.  In the discussions below, we'll use
the simulation names defined in the table to refer to specific runs
and we'll use the 10/10 run as the reference calculation.

\begin{deluxetable}{lcccccc}
\tablecaption{\label{table:sim_names} Simulation parameters.}
\tablehead{\colhead{name} &
           \colhead{reaction} &
           \colhead{conductivity} &
           \colhead{fine grid} &
           \colhead{rotation} &
           \colhead{domain size} &
           \colhead{network} \\
%
           \colhead{} &
           \colhead{boost} &
           \colhead{boost} &
           \colhead{resolution} &
           \colhead{rate} &
           \colhead{} &
           \colhead{}
}
\startdata
10/10 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ & {\tt aprox13} \\
10/10-slow & 10 & 10 & 10~cm & 1000~Hz & $1.8432\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
10/10-lores & 10 & 10 & 20~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt aprox13} \\
5/5 & 5 & 5 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$  &{\tt aprox13} \\
10/10-iso7 & 10 & 10 & 10~cm & 2000~Hz & $1.2288\times 10^5~\mbox{cm} \times 3.072\times 10^4~\mbox{cm}$ &{\tt iso7} \\
\enddata
\end{deluxetable}

\subsection{General Features}

Figure~\ref{fig:time_series} shows the time evolution of the 10/10
simulation, focusing on the mean-molecular weight,
\begin{equation}
\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}.
\end{equation}
In each frame the buffer of \isot{Ni}{56} that serves as the
underlying neutron star is seen spanning the bottom of the domain.
Above that, the composition begins as pure \isot{He}{4}, but as the
simulation progresses, the flame processes this to heavier nuclei,
increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
is reasonable well-defined.  We see that the bottom of the burning
front is lifted off of the base of the atmosphere, greatly increasing
the surface area of the burning compared with a perfectly vertical
flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
and we are beginning to see a gradient in the composition of the ash,
with the heavier nuclei furthest behind the flame.  The boosting of
the burning likely artificially increases this effect. \MarginPar{Doesn't the increased flame speed of the 10/10 run vs. the 5/5 run show this?}

% made with time_series.py in flame_wave_boost_10_10_paper with --skip 100
\begin{figure}[t]
\centering
\plotone{time_series}
\caption{\label{fig:time_series} Time series of the mean molecular weight of the 
flame for our standard 10/10 simulation.}
\end{figure}

Figure~\ref{fig:10_10_overview} shows the temperature, energy
generation rate, and $w$ component of the velocity (the out-of-plane
velocity induced by the Coriolis force).  This latter field
illustrates the hurricane effect set up by the laterally spreading
burning front.  In the energy generation rate plot, we see that the
burning is mostly concentrated toward the bottom of the layer, as
expected since the density is greatest there.  We see that the peak of
the burning has moved off of the symmetry axis, demonstrating that the
burning front is propagating to the right.  In the temperature plot,
we see the effect of our refinement criteria focusing only on the
atmosphere where we are most dense, with an artificial change in the
temperature at the refinement boundary due in part to the construction
of the initial model using the fine grid resolution for HSE. \MarginPar{Remind the reader what this
  was/where they should be looking in the plot to see this temperature
  drop} As we'll see when the lower resolution case, we do not believe
that this affects the results.

A final feature worth noting is the ash that seems to move along the
surface at a higher velocity than the flame, via surface gravity
waves.  The sponging that we perform is likely damping this to some extent, and
the method by which we initialize the flame may induce a larger
transient than in nature. Nevertheless, this surface ash is intriguing
because it affects the composition of the photosphere ahead of the
burning front, potentially changing our interpretation of
observations.  This is something that will be explored more fully in
the future.

% made with overview.py
\begin{figure}[t]
\centering
\plotone{flame_wave_boost_10_10_overview}
\caption{\label{fig:10_10_overview} Temperature, energy generation rate, and out-of-plane velocity for the 10/10 simulation at 20~ms.}
\end{figure}

Our default resolution puts $\sim 80$ zones vertically in the
``cool'' \MarginPar{double check} model atmosphere height.  To
understand the effects of resolution, we also performed a run with one
fewer refinement level, giving a $20~\mbox{cm}$ resolution overall (this is our
10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields for this run at
20~ms.  The structure is largely the same as the 10/10 run, with
largely the same flame shape and position, and the same structure in
the energy generation rate.  Since we do not have a jump in refinement
right below the atmosphere, we don't see the temperature feature from
the initial model mapping there, but we do see some cooling in the
\isot{Ni}{56} region.  This is likely a resolution effect, again due
to the strong gradient in temperature at the base of the atmosphere.
The strong agreement in the low resolution run to the 10/10 run gives
us confidence that we are capturing the flame physics properly.

% made with slice_multi_crop.py
\begin{figure}[t]
\centering
\plotone{flame_wave_boost_10_10_lowres_slice}
\caption{\label{fig:10_10_lowres} Temperature, mean molecular weight,
  energy generation rate, and out-of-plane velocity for the 10/10 low
  resolution simulation at 20~ms.}
\end{figure}

\subsection{Effects of our Approximations}

The results above all used a boosting of 10/10.  To see how the
results are sensitive to this boosting, we ran a simulation with a reduced
boosting of 5/5.  This is shown in Figure~\ref{fig:5_5_overview}.  The
results look qualitatively the same---a laterally propagating flame
develops that is lifted off of the bottom of the fuel layer.  The
flame has not moved out as far as in the 10/10 simulation, simply
because there is less energy release, but we expect that if we were to
run this out twice as long, the flame would have advanced to the
position seen in our 10/10 runs.  The good agreement in the structure
of the flame seen with the lower boosting gives us confidence that the
overall aspects of the flame structure and acceleration we are seeing
are robust to the approximations we make.

\begin{figure}[t]
\centering
\plotone{flame_wave_boost_5_5_slice}
\caption{\label{fig:5_5_overview} Temperature, mean molecular weight, energy generation rate, and out-of-plane velocity for the 5/5 simulation at 20~ms.}
\end{figure}

We also considered the effect of the network size.
Figure~\ref{fig:network} shows a comparison of the 10/10 boosting run
with the standard 13-isotope ``aprox13'' network and the reduced
7-isotope ``iso7'' network.  We see that the flame in the aprox13 case
is slightly more advanced and has a higher $\bar{A}$ than the iso7
case.  We see in the next section that these two networks give largely
the same flame speed.  The reduced network size saves a lot of memory,
which will be useful when we transition to 3-d simulations.

\begin{figure}[t]
\centering
\plotone{network_compare}
\caption{\label{fig:network} Mean molecular weight at 15~ms comparing a run with
  aprox13 vs a run with the iso7 network.}
\end{figure}

The final approximation to explore is our choice of rotation rate.
{\color{red} need to add discussion + plot}

\subsection{Flame Acceleration}

>>>>

\begin{document}
%======================================================================
% Title
%======================================================================
\title{Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning}

\shorttitle{Lateral Flame Dynamics II}
\shortauthors{Harpole et al.}

@@

\correspondingauthor{Alice Harpole}
\email{alice.harpolee@stonybrook.edu}

%======================================================================
% Abstract and Keywords
%======================================================================
\begin{abstract}
We continue to investigate laterally propagating flames in XRBs...
\end{abstract}

@@

@@

@@
